The Impact of Object Immutability on the Java Class Size

### 2020-04-19 paper/article.tex
MIT license
more

% \documentclass[a4paper,UKenglish,cleveref, autoref]{lipics-v2019}
\documentclass[12pt]{article}
\title{The Impact of Immutability on Class Size}
\author{Yegor Bugayenko}{}{}
% \authorrunning{Yegor Bugayenko}
% \ccsdesc[100]{Software and its engineering~Object oriented languages}
% \keywords{Object-Oriented Programming, Immutability, NCSS}
\usepackage[utf8]{inputenc}
\usepackage[numbers]{natbib}
\bibliographystyle{plainnat}
\usepackage[inline]{enumitem}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{verbatimbox}
\usepackage{interval}
\usepackage{hyperref}
\usepackage{minted}
  \setminted{fontsize=\footnotesize}
  \setminted{breaklines}
  \usemintedstyle{bw}
\newcommand{\code}[1]{\texttt{#1}}
\newenvironment{nicetable}
  {\setlength{\parindent}{0em}\medskip\small}
  {\medskip}
\input{total}
\begin{document}
\raggedbottom
\maketitle

\begin{abstract}
According to subjective opinions of many industry experts,
object immutability is a virtue in object-oriented programming, since it
leads to side-effect free design, cleaner code, better concurrency,
and many other factors. However, it has never been empirically
demonstrated exactly how immutability affects quality metrics of
object-oriented program. In the following research we analyzed
\thetotaljavafiles{} classes from \thetotalrepos{} public Java repositories
to find out how immutability affects the size of code.
\end{abstract}

@@

\section{Related Work}

@@

However, to our knowledge, the impact of immutability of class size has not been
analyzed yet.

\section{Empirical Results}

@@

\section{Conclusion}

@@

The source code of Ruby and Python scripts used to do the research
is available in GitHub repository \texttt{yegor256/size-vs-immutability}.

\bibliography{main}

>>>>

% The MIT License (MIT)
%
% Copyright (c) 2020 Yegor Bugayenko
%
% Permission is hereby granted, free of charge, to any person obtaining a copy
% of this software and associated documentation files (the "Software"), to deal
% in the Software without restriction, including without limitation the rights
% to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
% copies of the Software, and to permit persons to whom the Software is
% furnished to do so, subject to the following conditions:
%
% The above copyright notice and this permission notice shall be included
% in all copies or substantial portions of the Software.
%
% THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
% IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
% FITNESS FOR A PARTICULAR PURPOSE AND NON-INFRINGEMENT. IN NO EVENT SHALL THE
% AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
% LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
% OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
% SOFTWARE.

\documentclass[sigconf]{acmart}
\title{The Impact of Object Immutability on the Java Class Size}
\author{Yegor Bugayenko}{}{}
\email{yegor.bugayenko@huawei.com}
\affiliation{%
  \institution{Huawei Technologies Co., Ltd.}
  \city{Moscow, Russia}
}
\ccsdesc[100]{Software and its engineering~Object oriented languages}
\keywords{Object-Oriented Programming, Immutability, NCSS}
\setcopyright{acmcopyright}
\copyrightyear{2020}
\acmYear{2020}

\usepackage[utf8]{inputenc}
\usepackage[inline]{enumitem}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{verbatimbox}
\usepackage{interval}
\usepackage{hyperref}
\usepackage{minted}
  \setminted{fontsize=\footnotesize}
  \setminted{breaklines}
  \usemintedstyle{bw}
\newcommand{\code}[1]{\texttt{#1}}
\newenvironment{nicetable}
  {\setlength{\parindent}{0em}\medskip\small}
  {\medskip}
\input{total}
\begin{document}

\begin{abstract}
According to subjective opinions of many industry experts,
object immutability is a virtue in object-oriented programming, since it
leads to side-effect free design, cleaner code, better concurrency,
and many other factors. However, it has never been empirically
demonstrated exactly how immutability affects quality metrics of
object-oriented program. In the following research we analyzed
\thetotaljavafiles{} classes from \thetotalrepos{} public Java repositories
to find out how immutability affects the size of code.
\end{abstract}
\maketitle

@@

The paper is organized as follows.
Section~\ref{sec:background} defines various terms used in the paper.
Section~\ref{sec:related} covers related work in the areas of immutability and class size.
Section~\ref{sec:results} covers our empirical case study of \thetotalrepos{} Java source code repositories.
Section~\ref{sec:discussion} covers limitations of both the metric and the study.
Finally, we summarize our study in Section~\ref{sec:conclusion}.

\section{Background}
\label{sec:background}

Immutability in Java is a compile-time restriction of an object, making it
impossible for other objects to modify any of its attributes. For example,
objects of this class are immutable, thanks to the \texttt{final} modifier
attached to all of its attributes:

\begin{minted}{text}
class Book {
  private final int id;
  private final String title;
  Book(int i, String t) {
    this.id = i;
    this.title = t;
  }
}
\end{minted}

The only way to modify \texttt{final} attributes is through the constructor
of the class, which in Java by convention has the same name as the class (\text{Book}
in the example above).

A class of any size may be immutable. A class with no attributes is immutable.
A utility class, which is a class with no attributes, many static methods,
and a private constructor, is immutable.

\section{Related Work}
\label{sec:related}

@@

However, to our knowledge, the impact of immutability on the Java
class size has not been analyzed yet.

\section{Empirical Results}
\label{sec:results}

@@

\section{Discussion}
\label{sec:discussion}

To be written...

\section{Conclusion}
\label{sec:conclusion}

@@

% The source code of Ruby and Python scripts used to do the research
% is available in GitHub repository \texttt{yegor256/size-vs-immutability}.

\bibliographystyle{ACM-Reference-Format}
\bibliography{main}
